---
title: "Text Prediction Application"
author: "Lucas Qualmann"
date: "11/29/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Text Prediction Application

- Can you never remember the next word you want to say?
- Are you worried your text prediction keyboards are storing and selling your data?
- Do you like learning new words?

This app is a perfect fit for you!

## How It Works

- Type a phrase in the text box ommitting the word you want predicted and hit the submit button.
- You will see the application's first, second, and third choice of predicted words.
- Compare it to the word you were thinking of using.

It's that simple!

[Here's the link to the application!](https://lqual.shinyapps.io/Shiny_App/)

## Example

- Phrase: "Let's go to the"
- Our next word: "mall"
- Predictions:

```{r, echo=FALSE, message=FALSE, comment=""}
source('~/Capstone_Project/Model/textPredictor.R')
answer <- getword("Let's go to the")
paste("Prediction1: ", answer[1, 1], sep = " ")
paste("Prediction2: ", answer[2, 1], sep = " ")
paste("Prediction3: ", answer[3, 1], sep = " ")
```

Our next word was the second prediction of the application!

## Data Used for Application

- The application uses ngrams as large as 7 words to as small as 1 word to predict the next word.
- If the application comes accross an unkown phrase, it selects 3 of the 10 most observed words in our dataset according to their number of occurences: the, and, for, that, you, with, was, this, have, or are.
- Example of unknown phrase output below.
```{r, echo=FALSE, message=FALSE, comment=""}
myprobs <- c(10/26, 5/26, 2/26, 2/26, 2/26, 1/26, 1/26, 1/26, 1/26, 1/26)
lookup1 <- sample(1:10, size = 3, replace = FALSE, prob = myprobs)
unigram <- data.frame(next_word = c("the", "and", "for", "that", "you",
                                    "with", "was", "this", "have","are"))
answer1 <- data.frame(unigram$next_word[lookup1])
colnames(answer1) <- "answer"
paste("Prediction1: ", answer1[1, 1], sep = " ")
paste("Prediction2: ", answer1[2, 1], sep = " ")
paste("Prediction3: ", answer1[3, 1], sep = " ")
```










